Fourmile Calibrations and Output

### 2025-04-09 fourmile_output_dec2024.Rmd
redid kge and pbias plot and added nse

@@

flow_plot_updated <- flow_plot_updated %>%
  mutate(kge_revise = if_else(kge < 0, 0, kge))

@@

flow_plot_updated <- flow_plot_updated %>%
  mutate(kge_revise_bins = cut(kge,
                               breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, 1),
                               include.lowest = TRUE,
                               labels = c("< 0.0", "0.2-0.4", "0.4-0.6", "0.6-0.8", "> 0.8")))

@@

both_kge_plot <- flow_plot_updated %>%
  ggplot(., aes(x = year_excluded, 
              y = year, 
              fill = kge_revise_bins)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  paletteer::scale_fill_paletteer_d("PNWColors::Anemone") +
  geom_text(aes(label = sprintf("%.2f", 
                                kge)), 
            color = "black", 
            size = 3) +
  theme_bw(base_size=18) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  labs(y = "Year",
       x='Year excluded',
       fill = 'KGE') +
  theme(
    axis.text.x = element_text(angle = 45, 
                               hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(.75, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt"),
    # axis.title.y = element_blank(),
    # axis.text.y = element_blank(),
    # axis.ticks.y = element_blank()
  )

@@

#####################
kgeCats <- flow_plot_updated %>%
  mutate(pbiasCat = case_when(
    abs(pbias) < 10 ~ 'Very good',
    abs(pbias) >= 10 & abs(pbias) < 15 ~ 'Good',
    abs(pbias) >= 15 & abs(pbias) < 25 ~ 'Satisfactory',
    abs(pbias) >= 25 ~ 'Unsatisfactory'),
    pbiasCat = factor(pbiasCat, levels = c("Unsatisfactory", 
                                           "Satisfactory", 
                                           "Good", 
                                           "Very good")))

both_pbias_plot <- pbiasCats %>%
  ggplot(aes(x = year_excluded, 
             y = year, 
             fill = pbiasCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  # paletteer::scale_fill_paletteer_c(
  #   "grDevices::Green-Brown",
  #   limits = c(-max_abs, max_abs),
  #   oob = scales::squish
  # ) +
  paletteer::scale_fill_paletteer_d("calecopal::lupinus") +
  geom_text(aes(label = sprintf("%.2f", pbias)), 
            color = "black", 
            size = 2.75) +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year",
       x = 'Year excluded',
       fill = '% bias')

@@

both_pbias_plot <- pbiasCats %>%
  ggplot(aes(x = year_excluded, 
             y = year, 
             fill = pbiasCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  # paletteer::scale_fill_paletteer_c(
  #   "grDevices::Green-Brown",
  #   limits = c(-max_abs, max_abs),
  #   oob = scales::squish
  # ) +
  paletteer::scale_fill_paletteer_d("calecopal::lupinus") +
  #paletteer::scale_fill_paletteer_d("PNWColors::Anemone") +
  geom_text(aes(label = sprintf("%.2f", pbias)), 
            color = "black", 
            size = 2.75) +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year",
       x = 'Year excluded',
       fill = '% bias')

@@

# max_abs <- max(abs(flow_plot_updated$pbias), na.rm = TRUE)
# 
# both_pbias_plot <- flow_plot_updated %>%
#   ggplot(aes(x = year_excluded, 
#              y = year, 
#              fill = pbias)) +
#   geom_tile(color = "white") +
#   facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
#   paletteer::scale_fill_paletteer_c(
#     "grDevices::Green-Brown",
#     limits = c(-max_abs, max_abs),
#     oob = scales::squish
#   ) +
#   geom_text(aes(label = sprintf("%.2f", pbias)), 
#             color = "black", 
#             size = 2.75) +
#   theme_bw(base_size=18) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.position = 'bottom',
#     legend.text = element_text(size = 14),
#     legend.title = element_text(size = 14),
#     legend.key.width = unit(2, "cm"),
#     panel.grid = element_blank(),
#     panel.background = element_rect(fill = "white"),
#     strip.background = element_blank(),
#     text = element_text(color = "black"),
#     axis.title = element_text(color = "black"),
#     axis.text = element_text(color = "black"),
#     plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
#     legend.margin = margin(t = -8, b = -4, unit = "pt")
#   ) +
#   labs(y = "Year",
#        x = 'Year excluded',
#        fill = 'Pbias')
# 
# both_pbias_plot

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/pbias_heatmap_both_basic_0center.png',
       dpi=600,
       width = 10,
       height = 6)

>>>>

## NSE

flow_nse_all <- flow %>%
  drop_na() %>%
  summarize(grid2011 = NSE(grid_2011, obs),
            grid2012 = NSE(grid_2012, obs),
            grid2013 = NSE(grid_2013, obs),
            grid2014 = NSE(grid_2014, obs),
            grid2015 = NSE(grid_2015, obs),
            grid2016 = NSE(grid_2016, obs),
            grid2017 = NSE(grid_2017, obs),
            grid2018 = NSE(grid_2018, obs),
            grid2019 = NSE(grid_2019, obs),
            grid2020 = NSE(grid_2020, obs),
            gridall = NSE(grid_all, obs),
            statn2011 = NSE(statn_2011, obs),
            statn2012 = NSE(statn_2012, obs),
            statn2013 = NSE(statn_2013, obs),
            statn2014 = NSE(statn_2014, obs),
            statn2015 = NSE(statn_2015, obs),
            statn2016 = NSE(statn_2016, obs),
            statn2017 = NSE(statn_2017, obs),
            statn2018 = NSE(statn_2018, obs),
            statn2019 = NSE(statn_2019, obs),
            statn2020 = NSE(statn_2020, obs),
            statnall = NSE(statn_all, obs)) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_nse_all <- flow_nse_all %>%
  pivot_longer(everything()) %>%
  arrange(value)

# determine nse for each year for every model run
flow_year <- flow %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(
    grid2011 = NSE(grid_2011, obs),
            grid2012 = NSE(grid_2012, obs),
            grid2013 = NSE(grid_2013, obs),
            grid2014 = NSE(grid_2014, obs),
            grid2015 = NSE(grid_2015, obs),
            grid2016 = NSE(grid_2016, obs),
            grid2017 = NSE(grid_2017, obs),
            grid2018 = NSE(grid_2018, obs),
            grid2019 = NSE(grid_2019, obs),
            grid2020 = NSE(grid_2020, obs),
            gridall = NSE(grid_all, obs, na.rm = T, method='2012'),
            statn2011 = NSE(statn_2011, obs),
            statn2012 = NSE(statn_2012, obs),
            statn2013 = NSE(statn_2013, obs),
            statn2014 = NSE(statn_2014, obs),
            statn2015 = NSE(statn_2015, obs),
            statn2016 = NSE(statn_2016, obs),
            statn2017 = NSE(statn_2017, obs),
            statn2018 = NSE(statn_2018, obs),
            statn2019 = NSE(statn_2019, obs),
            statn2020 = NSE(statn_2020, obs),
            statnall = NSE(statn_all, obs, na.rm=T, method='2012')) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)

flow_nse <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>%
  pivot_wider(names_from = 'year', values_from = 'nse')

flow_plot <- flow_year %>%
  pivot_longer(!year, names_to = 'model_run', values_to = 'nse') %>%
  mutate(year = as.factor(year))

# add all to the plot below
nse_all <- flow_nse_all %>%
  mutate(year = 'All') %>%
  rename(model_run = name,
         nse = value)

flow_plot_updated <- bind_rows(flow_plot, nse_all)

# nse figure
flow_plot_updated <- flow_plot_updated %>% 
  mutate(source = case_when(
    grepl("grid", model_run) ~ "grid",
    grepl("statn", model_run) ~ "statn"),
    year_excluded = str_extract(model_run, "\\d{4}$")) %>%
  mutate(year_excluded = if_else(model_run %in% c('gridall',
                                                       'statnall'),
                                  'None', year_excluded))

# put nse into categories
nseCats <- flow_plot_updated %>%
  mutate(nseCat = case_when(
    nse > 0.75 ~ 'Very good',
    nse > 0.65 & nse <= 0.75 ~ 'Good',
    nse > 0.5 & nse <= 0.65 ~ 'Satisfactory',
    nse <= 0.5  ~ 'Unsatisfactory'),
    nseCat = factor(nseCat, levels = c("Unsatisfactory", 
                                           "Satisfactory", 
                                           "Good", 
                                           "Very good")))

write_csv(nseCats, '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/nseCats.csv')

both_nse_plot <- nseCats %>%
  ggplot(aes(x = year, 
             y = year_excluded, 
             fill = nseCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  scale_fill_manual(values = c(
    "Unsatisfactory" = "#F1F4EEFF", 
    "Satisfactory" = "#EFDDCFFF",
    "Good" = "#72E1E1FF",
    "Very good" = "#009474FF"
  )) +
  geom_text(aes(label = sprintf("%.2f", nse)), 
            color = "black", 
            size = 2.75) +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year excluded",
       x = 'Year',
       fill = 'NSE')

both_nse_plot

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/nse_heatmap_both_cats.png',
       dpi=600,
       width = 10,
       height = 6)

@@

# flow_plot_updated <- flow_plot_updated %>%
#   mutate(kge_revise = if_else(kge < 0, 0, kge))

@@

flow_plot_updated <- flow_plot_updated %>%
  mutate(kge_revise_bins = cut(kge,
                               breaks = c(-Inf, 0.0, 0.2, 0.4, 0.6, 0.8, 1),
                               include.lowest = TRUE,
                               labels = c("< 0.0","0.0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "> 0.8")))

@@

both_kge_plot <- flow_plot_updated %>%
  ggplot(., aes(x = year, 
              y = year_excluded, 
              fill = kge_revise_bins)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  #paletteer::scale_fill_paletteer_d("PNWColors::Anemone") +
  # paletteer::scale_fill_paletteer_d(
  # "PNWColors::Anemone", 
  # guide = guide_legend(nrow = 1),
  # direction = -1) +
   scale_fill_manual(values = c(
    "< 0.0" = "#DCBE9BFF", 
    "0.0-0.2" = "#EFDDCFFF",
    "0.2-0.4" = "#F1F4EEFF",
    "0.4-0.6" = "#72E1E1FF",
    "0.6-0.8" = "#11C2B5FF",
    "> 0.8" = "#009474FF"),
    guide = guide_legend(nrow = 1)) +
  geom_text(aes(label = sprintf("%.2f", 
                                kge)), 
            color = "black", 
            size = 3) +
  theme_bw(base_size=18) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  labs(y = "Year excluded",
       x='Year',
       fill = 'KGE') +
  theme(
    axis.text.x = element_text(angle = 45, 
                               hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(.6, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt"),
  )

@@

@@

both_pbias_plot <- pbiasCats %>%
  ggplot(aes(x = year, 
             y = year_excluded, 
             fill = pbiasCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
   scale_fill_manual(values = c(
    "Unsatisfactory" = "#F1F4EEFF", 
    "Satisfactory" = "#EFDDCFFF",
    "Good" = "#72E1E1FF",
    "Very good" = "#009474FF"
  )) +
  geom_text(aes(label = sprintf("%.2f", pbias)), 
            color = "black", 
            size = 2.75) +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year excluded",
       x = 'Year',
       fill = '% bias')

@@
